Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-07-12 vesicle_code/docs/adhesion/adhesion.tex
yny revision July 12

  \includegraphics[height = 0.2\textwidth,trim={1cm 0cm 1cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image01.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image02.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image03.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image04.png}
  \includegraphics[height = 0.2\textwidth,trim={3.5cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image05.png}
  \caption{\label{fig:extensionalNonsymmetric}Dynamics of a vesicle pair in a planar extensional flow. ${\cal H} = 0.7$ for the first three rows, and ${\cal H}=0.6$ for the bottom row. From top to bottom, the initial vertical displacement $\triangle y=0$, 0.02, 0.1 and 0.1. The reduced area is ???, the extensional rate $\chi = ???$ and the adhesion distance $\delta = ???$. }
  \end{figure}

The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  Here we focus on a
planar extensional flow  where the fluid velocity field $\uu=\chi(-x,y)$ is centered at the 
the origin ($(x,y)=(0,0)$), the stagnation point where the fluid flow converges horizontally and diverges vertically.
Placed symmetrically at the stagnation point of an extensional flow,
a single vesicle may remain steady and symmetric for low extensional rate $\chi$.
For sufficiently large extensional rate $\chi$ the vesicle may undergo asymmetric deformation \cite{KantslerSegreSteinberg2008_PRL,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt}.

To investigate how two vesicles interact with each other in a planar extensional flow,
at the beginning of the simulations
we place a vesicle at the stagnation point (left panels in Figure~\ref{fig:extensionalNonsymmetric}), 
and the other vesicle at $(x,y) = (3,\triangle y)$ where the initial vertical displacement $\triangle y$ varies from
$\triangle y=0$ for the top row, $\triangle y = 0.02$ for the second row, to
$\triangle y =0.1$ for the bottom two rows in Figure~\ref{fig:extensionalNonsymmetric}.
The adhesion strength ${\cal H} = 0.7$ for the top three rows, and ${\cal H} =0.6$ for the bottom row.

Once the planar extensional flow is turned on at $t=0^+$, two vesicles move
towards each other  as shown in  Figure \ref{fig:extensionalNonsymmetric}.
For $\triangle y=0$ in the top row, the two vesicles stay on the $x$-axis as they move closer to 
form a doublet. We observe that the left vesicle (initially centered at $(0,0)$) moves to the left due to the impinging
vesicle from the right. As time progresses they form a doublet
and reach a steady configuration that is symmetric with respect to the stagnation point 
(filled squares in Figure \ref{fig:extensionalNonsymmetric}).

As the initial vertical displacement $\triangle y$ increases, we observe similar vesicle dynamics only at the beginning 
when the right vesicle moves towards the left vesicle as the left vesicle is pushed to the left of the stagnation point. 
As both vesicles come to close vicinity of the stagnation point, the diverging flow pulls the
right vesicle (elevated with a small vertical displacement at $t=0$) further away, causing rotation of both vesicles. 
We also observe that the larger the initial vertical displacement, the larger the rotation and the faster both vesicles move
away from the stagnation point.

For sufficiently strong adhesion (such as ${\cal H}=0.7$ for first three rows in Figure \ref{fig:extensionalNonsymmetric}),
 the two vesicles rotate and move away from the stagnation point as a doublet until the end of simulations ($t=100$).
As the two vesicles move further away from the stagnation point, the extensional flow becomes more effective in pulling the two drops apart
as shown in the third row of Figure \ref{fig:extensionalNonsymmetric}.
For ${\cal H}=0.6$ (bottom row in Figure \ref{fig:extensionalNonsymmetric})  
the two vesicles are separated by the diverging flow as early as $t=75$, and are clearly not a doublet at the end of simulation.

Results in Figure~\ref{fig:extensionalNonsymmetric} illustrate that
in an extensional flow vesicles are prone to move away from the stagnation point:
A small initial vertical displacement $\triangle y$ in one  of the vesicles causes both vesicles to not only rotate but also move away from the
center. Once vesicles move along the diverging stream, their hydrodynamics is dominated by the strong straining flow in the far-field.
Thus in the microfluidic experiment it is often more desirable to trap the vesicle around the center of an extensional flow \cite{Spjut2010_MSThesis_Chapter3}.

With an active
control algorithm to place the stagnation point at a desirable location
by adjusting the streaming flow strength with a feedback
loop~\cite{Johnson-Chavarria2011_EMJ}, a particle can be trapped at the
stagnation point for long time scales to facilitate image acquisition or
other detailed measures such as particle image velocimetry of flow
inside and around the particle.  To explore the application of such a
fluid trap to measure the adhesion strength between two bound vesicle
membranes, we propose the following experiment.  Beginning with two
identical vesicles in an equilibrium configuration that form a doublet
with a flat contact region (Figure~\ref{fig:Dec18_vesicle_shape}), we
turn on the fluid trap with the stagnation point placed at the center of
the vesicle doublet.  Depending on the Hamaker constant, we expect that
either the flow overcomes the adhesive force and the doublet is broken,
or the adhesive force is sufficiently strong and the doublet reaches a
stable stationary configuration.

%In the following numerical experiments we initially place a vesicle
%doublet with $\theta=\pi$, so that the diverging flow may be strong
%enough to pull vesicles away from the doublet.  At low extensional
%rates, we expect the doublet to stay bound at the fluid trap stagnation
%point.  On the other hand, the vesicle doublet may become unstable and
%eventually separate at higher extensional rates.  Thus we expect there
%to exist a critical extensional rate $\chi_c$ above which the vesicle
%doublet cannot stay bound under a given adhesion potential.  Therefore,
%the dependence of the critical extensional rate $\chi_c$ on the adhesion
%potential and the mechanical vesicle properties provides a means to
%probe the physics of membrane adhesion.

@@

Based on how the contact region aligns with the extensional flow, we can
define the inclination angle of the vesicle doublet as illustrated in
Figure~\ref{fig:InclinationAngle}. When $\theta=\pi$, the doublet long
axis is aligned with the far-field converging flow, and the diverging flow
pulls the vesicles apart from each other and the long-range attractive
force is essential to keep the two vesicles from separating.  In
contrast, when $\theta = \pi/2$, the doublet long axis is aligned with
the diverging flow, the converging flow pushes the vesicles towards
each other and the short-range repulsive force is essential to keep the
two vesicles at the separation distance.  With the converging flow
pushing the two vesicles towards the stagnation point dynamically placed
at the doublet center, this configuration is expected to be more stable
than the $\theta=\pi$ configuration in the left of
Figure~\ref{fig:InclinationAngle}(a). This is because when $\theta=\pi$
the diverging flow is pulling the vesicles away from each other as shown
in the illustration.  However, when the doublet inclination angle is
fixed at $\theta=\pi/2$, the vesicle doublet stays intact if the
adhesive force is strong enough to dominate the fluid drag.  However, it
is possible that the diverging flow may overcome adhesion and break up
the doublet.  It is also reasonable to expect a vesicle doublet to
rotate from the less stable configuration $\theta=\pi$ to the more
stable configuration $\theta=\pi/2$, so that the fluid flow is pushing
them together rather than pulling them apart.

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}\\
   \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png} \\
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional1} A vesicle doublet in an extensional
  flow with an initial inclination angle $\theta=\pi$.  The reduced
  area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} = 0.7$,
  the separation distance is $\delta = 0.4$, and the extensional rate is
  $\chi = 0.02$ ({\em top}), $\chi=0.07$ ({\em middle}), and $\chi =
  0.1$ ({\em bottom}).  The center of the vesicle, denoted with a black
  square, is the stagnation point.}
  \end{figure}

In the following numerical experiments we initially place a vesicle
doublet with $\theta=\pi$, so that the diverging flow may be strong
enough to pull vesicles away from the doublet.  At low extensional
rates, we expect the doublet to stay bound at the fluid trap stagnation
point.  On the other hand, the vesicle doublet may become unstable and
eventually separate at higher extensional rates.  Thus we expect there
to exist a critical extensional rate $\chi_c$ above which the vesicle
doublet cannot stay bound under a given adhesion potential.  Therefore,
the dependence of the critical extensional rate $\chi_c$ on the adhesion
potential and the mechanical vesicle properties provides a means to
probe the physics of membrane adhesion.

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $2\pi$, and we vary the extensional
rate, $\chi$, between $10^{-2}$ and $10^{-1}$.  Since the stagnation
point can be controlled in an experimental
setting~\cite{Johnson-Chavarria2011_EMJ}, we mimic the active control of
the microfluidic experiments by moving the center of the doublet at each
time step so that the stagnation point occurs exactly in the middle of
the doublet.  With this adjustment, the vesicles either remain as a
doublet in the fluid trap centered around the stagnation point, or the
doublet is broken and the vesicles separate from one another.

In Figures~\ref{fig:extensional1}, snapshots of a fluid trap containing a
vesicle doublet of reduced area $\Delta A = 0.7$ are simulated at three
different extensional rates.  We observe that the doublet rotates from
$\theta=\pi$ configuration towards the more stable $\theta=\pi/2$
configuration, as shown in Figures~\ref{fig:extensional1}.  
%
The final configurations in Figures~\ref{fig:extensional1}
are similar to those at $t=100$ in Figure~\ref{fig:extensionalNonsymmetric}.
%
We also
observe that with moderate extensional rates, the doublet falls short of
aligning their long axis with the diverging direction.  Moreover, the
final inclination angle is closest to the stable $\theta=\pi/2$ for the
smallest extensional rates. While this doublet remains bound for all
the considered extensional rates, a doublet with a reduced area of
$\Delta A = 0.8$ is split with a critical extensional rate $\chi_c =
0.1$ (Figure~\ref{fig:extensional3}).

%\begin{figure}[htp]
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}\\
%   \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png} \\
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
%  \caption{\label{fig:extensional1} A vesicle doublet in an extensional
%  flow with an initial inclination angle $\theta=\pi$.  The reduced
%  area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} = 0.7$,
%  the separation distance is $\delta = 0.4$, and the extensional rate is
%  $\chi = 0.02$ ({\em top}), $\chi=0.07$ ({\em middle}), and $\chi =
%  0.1$ ({\em bottom}).  The center of the vesicle, denoted with a black
%  square, is the stagnation point.}
%  \end{figure}

>>>>

  \includegraphics[height = 0.2\textwidth,trim={1cm 0cm 1cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image01.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image02.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image03.png}
  \includegraphics[height = 0.2\textwidth,trim={4cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image04.png}
  \includegraphics[height = 0.2\textwidth,trim={3.5cm 0cm 4cm 0cm},clip]
    {figs/extensional_displacement1em1_adS6em1_image05.png}
  \caption{\label{fig:extensionalNonsymmetric}Dynamics of a vesicle pair in a planar extensional flow. ${\cal H} = 0.7$ for the first three rows, and ${\cal H}=0.6$ for the bottom row. From top to bottom, the initial vertical displacement $\triangle y=0$, 0.02, 0.1 and 0.1. The reduced area is ???, the extension rate $\chi = ???$ and the adhesion distance $\delta = ???$. }
  \end{figure}

The hydrodynamics of a single vesicle in an extensional flow has
revealed novel nonlinear vesicle dynamics not found for a viscous
drop~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}.  Here we focus on a
planar extensional flow  where the fluid velocity field $\uu=\chi(-x,y)$ is centered at the 
the origin ($(x,y)=(0,0)$), the stagnation point where the fluid flow converges horizontally and diverges vertically.
Placed symmetrically at the stagnation point of an extensional flow,
a single vesicle may remain steady and symmetric for low extension rate $\chi$.
For sufficiently large extension rate $\chi$ the vesicle may undergo asymmetric deformation \cite{KantslerSegreSteinberg2008_PRL,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt}, 
similar to a surfactant-laden viscous drop under a planar extensional flow \cite{JanssenBoonAgterof1997_AIChE,HuPineLeal2000_PoF}.

To investigate how two vesicles interact with each other in a planar extensional flow,
at the beginning of the simulations (left panels in Figure~\ref{fig:extensionalNonsymmetric})
we place a vesicle at $(x,y)=(0,0)$, 
and the other vesicle at $(x,y) = (3,\triangle y)$ where the initial vertical displacement $\triangle y$ varies from
$\triangle y=0$ for the top row, $\triangle y = 0.02$ for the second row, to
$\triangle y =0.1$ for the bottom two rows in Figure~\ref{fig:extensionalNonsymmetric}.
The adhesion strength ${\cal H} = 0.7$ for the top three rows, and ${\cal H} =0.6$ for the bottom row.

Once the planar extensional flow is turned on at $t=0^+$, two vesicles move
towards each other  as shown in  Figure \ref{fig:extensionalNonsymmetric}.
For $\triangle y=0$ in the top row, the two vesicles stay on the $x$-axis as they move closer to 
form a doublet. We observe that the left vesicle (initially centered at $(0,0)$) moves to the left due to the impinging
vesicle from the right. As time progresses they form a doublet
and reach a steady configuration that is symmetric with respect to the stagnation point 
(square mark in Figure \ref{fig:extensionalNonsymmetric}).

For $\triangle y>0$ we observe similar vesicle dynamics only at the beginning 
when the right vesicle moves towards the stagnation point and the left vesicle is pushed to the left of the stagnation point. 
As both vesicles come to close vicinity of the stagnation point, the elevated right vesicle ($\triangle y >0$) is pulled up by the diverging flow in the $y$ direction,
causing rotation of both vesicles. 
We also observe that the larger the initial vertical displacement, the larger the rotation and the faster both vesicles move
away from the stagnation point.
%
For sufficiently strong adhesion (${\cal H}=0.7$ for the first three rows in Figure \ref{fig:extensionalNonsymmetric}),
 the two vesicles rotate and move away from the stagnation point as a doublet until the end of simulations ($t=100$).
As the two vesicles move further away from the stagnation point, the extensional flow becomes more effective in pulling the two drops apart
as shown in the third row of Figure \ref{fig:extensionalNonsymmetric}.
For ${\cal H}=0.6$ (bottom row in Figure \ref{fig:extensionalNonsymmetric})  
the two vesicles are separated by the diverging flow as early as $t=75$, and are clearly not a doublet at the end of simulation.

Results in Figure~\ref{fig:extensionalNonsymmetric} illustrate that
in an extensional flow vesicles are prone to move away from the stagnation point:
A small initial vertical displacement $\triangle y$ in one  of the vesicles causes both vesicles to not only rotate but also move away from the
center. Once vesicles move off the horizontal axis, their hydrodynamics is dominated by the diverging flow as they move further away.
Thus in the microfluidic experiment it is often more desirable to trap the vesicle around the center of an extensional flow \cite{Spjut2010_MSThesis_Chapter3}.
%
With an active
control algorithm to place the stagnation point at a desirable location
by adjusting the flow with a feedback
loop~\cite{BentleyLeal1986_JFMa,Johnson-Chavarria2011_EMJ}, a particle can be trapped at the
stagnation point for long time scales to facilitate image acquisition or
other detailed measures such as particle image velocimetry of flow
inside and around the particle.

To explore the application of such a
fluid trap to measure the adhesion strength between two bound vesicles, 
we propose the following experiment.  
Beginning with two
identical vesicles in an equilibrium configuration that form a doublet
with a flat contact region (Figure~\ref{fig:Dec18_vesicle_shape}), we
turn on the fluid trap with the stagnation point placed at the center of
the vesicle doublet.  Depending on the Hamaker constant, we expect that
either the flow overcomes the adhesive force and the doublet is broken,
or the adhesive force is sufficiently strong and the doublet reaches a
stable stationary configuration.

%In the following numerical experiments we initially place a vesicle
%doublet with $\theta=\pi$, so that the diverging flow may be strong
%enough to pull vesicles away from the doublet.  At low extensional
%rates, we expect the doublet to stay bound at the fluid trap stagnation
%point.  On the other hand, the vesicle doublet may become unstable and
%eventually separate at higher extension rates.  Thus we expect there
%to exist a critical extension rate $\chi_c$ above which the vesicle
%doublet cannot stay bound under a given adhesion potential.  Therefore,
%the dependence of the critical extension rate $\chi_c$ on the adhesion
%potential and the mechanical vesicle properties provides a means to
%probe the physics of membrane adhesion.

@@

Based on how the contact region aligns with the extensional flow, we can
define the inclination angle $\theta$ of the vesicle doublet as illustrated in
Figure~\ref{fig:InclinationAngle}(b). When $\theta=\pi$ (Figure~\ref{fig:InclinationAngle}(a))
the diverging flow
pulls the vesicles apart from each other and the attractive adhesion
force is essential to keep the two vesicles from separating.  In
contrast, when $\theta = \pi/2$ (Figure~\ref{fig:InclinationAngle}(c)), 
%the doublet long axis is aligned with the diverging flow, 
the converging flow pushes the vesicles towards
each other and the repulsive force is essential to keep the
two vesicles at the separation distance.  With the converging flow
pushing the two vesicles towards the stagnation point dynamically placed
at the doublet center, we expect the $\theta=\pi/2$ configuration to be more stable
than the $\theta=\pi$ configuration. 
%
%This is because when $\theta=\pi$
%the diverging flow is pulling the vesicles away from each other as shown
%in the illustration.  However, when the doublet inclination angle is
%fixed at $\theta=\pi/2$, the vesicle doublet stays intact if the
%adhesive force is strong enough to dominate the fluid drag.  However, it
%is possible that the diverging flow may overcome adhesion and break up
%the doublet.  It is also reasonable to expect a vesicle doublet to
%rotate from the less stable configuration $\theta=\pi$ to the more
%stable configuration $\theta=\pi/2$, so that the fluid flow is pushing
%them together rather than pulling them apart.

\begin{figure}[htp]
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}\\
   \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png} \\
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
  \caption{\label{fig:extensional1} A vesicle doublet in an extensional
  flow with an initial inclination angle $\theta=\pi$.  The reduced
  area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} = 0.7$,
  the separation distance is $\delta = 0.4$, and the extension rate is
  $\chi = 0.02$ ({\em top}), $\chi=0.07$ ({\em middle}), and $\chi =
  0.1$ ({\em bottom}).  The center of the vesicle, denoted with a black
  square, is the stagnation point.}
  \end{figure}

In the following numerical experiments we initially place a vesicle
doublet with $\theta=\pi$, so that the diverging flow may be strong
enough to pull vesicles away from the doublet.  At low extension
rates, we expect the doublet to stay bound at the fluid trap stagnation
point.  On the other hand, the vesicle doublet may become unstable and
eventually separate at higher extension rates.  Thus we expect there
to exist a critical extension rate $\chi_c$ above which the vesicle
doublet cannot stay bound under a given adhesion potential.  Therefore,
the dependence of the critical extension rate $\chi_c$ on the adhesion
potential and the mechanical vesicle properties provides a means to
probe the physics of membrane adhesion.

We consider vesicle doublets with reduced areas 0.70, 0.75, 0.80, 0.85,
0.90, and 0.95, all with a length of $2\pi$, and we vary the extension
rate ($\chi$) between $10^{-2}$ and $10^{-1}$.  Since the stagnation
point can be controlled in an experimental
setting~\cite{BentleyLeal1986_JFMa,Johnson-Chavarria2011_EMJ}, we mimic the active control of
the microfluidic experiments by moving the center of the doublet at each
time step so that the stagnation point occurs exactly in the middle of
the doublet.  With this adjustment, the vesicles either remain as a
doublet in the fluid trap centered around the stagnation point, or the
doublet is broken and the vesicles separate from one another.

Figure~\ref{fig:extensional1} shows snapshots from simulating a vesicle doublet of reduced area  $\Delta A = 0.7$
at three different extension rates. A general feature of the doublet dynamics is that the doublet rotates from
$\theta=\pi$ configuration towards the more stable $\theta=\pi/2$
configuration.  
%
We also
observe that with moderate extension rates, the doublet falls short of
aligning their long axis with the diverging direction.  Moreover, the
final inclination angle is closest to the stable $\theta=\pi/2$ for the
smallest extension rates. While this doublet remains bound for all
the considered extension rates, a doublet with a reduced area of
$\Delta A = 0.8$ is split with a critical extension rate $\chi_c =
0.1$ (Figure~\ref{fig:extensional3}).

%\begin{figure}[htp]
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi2em2_ra070_image05.png}\\
%   \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi7em2_ra070_image05.png} \\
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image01.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image02.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image03.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image04.png}
%  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra070_image05.png}
%  \caption{\label{fig:extensional1} A vesicle doublet in an extensional
%  flow with an initial inclination angle $\theta=\pi$.  The reduced
%  area is $\Delta A = 0.7$, the Hamaker constant is $\mathcal{H} = 0.7$,
%  the separation distance is $\delta = 0.4$, and the extension rate is
%  $\chi = 0.02$ ({\em top}), $\chi=0.07$ ({\em middle}), and $\chi =
%  0.1$ ({\em bottom}).  The center of the vesicle, denoted with a black
%  square, is the stagnation point.}
%  \end{figure}

 \begin{figure}
    \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image01.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image02.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image03.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image04.png}
  \includegraphics[width = 0.16\textwidth,trim={4cm 2cm 4cm 1cm},clip]{figs/extensional_adR4em1adS7em1Chi1em1_ra080_image05.png}
  \caption{\label{fig:extensional3} A vesicle doublet in an extensional
  flow with an initial inclination angle $\theta=\pi$.  The reduced area
  is $\Delta A = 0.8$, the Hamaker constant $\mathcal{H} = 0.7$, the
  separation distance is $\delta = 0.4$, and the extension rate $\chi
  = 0.1$.  The center of the vesicle, denoted with a black square, is a
  stagnation point.}
\end{figure}

%
The equilibrium configuration for low extension rate $\chi=0.02$ (top row in Figure~\ref{fig:extensional1}) is
almost identical to the equilibrium configuration for $\triangle y=0$ (top row of  Figure~\ref{fig:extensionalNonsymmetric}).
The rotation of the vesicle doublet 
from $\theta=\pi$ to $\theta=\pi/2$ can be understood as the transition from an unstable configuration $\theta=\pi$ to a stable configuration $\theta=\pi/2$.
For higher extension rates (second and third rows in Figure~\ref{fig:extensional1}) the partially rotated vesicle pairs at equilibrium are
also very similar to the case of $\triangle y=0.02$ (second row of Figure~\ref{fig:extensionalNonsymmetric}).
Such asymmetric vesicle configuration is reminiscent of the asymmetric deformation of a single vesicle in an extensional flow \cite{KantslerSegreSteinberg2008_PRL,Narsimhan2014_JFM,DahlNarsimhanGouveia2016_SoftMatt}.
To gain more insight into such nonlinear dynamics, in the following we explore more physical characterization of the transition from $\theta=\pi$ to $\theta=\pi/2$ at
different reduced area and extension rate.
%
